---
title: "TidyTemplate"
date: 2022-03-01
output: html_document
editor_options: 
  chunk_output_type: console
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(rvest)
library(maps)
library(leaflet)

setwd('/Users/SG/Documents/Programming/TidyTuesday')
url = 'https://statesymbolsusa.org/symbol-official-item/national-us/uncategorized/states-size'
url_state = "https://raw.githubusercontent.com/vincentarelbundock/countrycode/master/data/custom_dictionaries/us_states.csv"

state_dict = read.csv(url_state, stringsAsFactors=FALSE)
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2022-03-01")

```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

tt

```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

tt %>% 
  map(glimpse)

```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}
alt_fuel <- tt$stations
'%notin%' <- Negate('%in%')

# Select relevant columns and public access
alt_fuel <-  alt_fuel %>%
  janitor::clean_names() %>%
  filter(access_code %in% 'public', state %notin% c('PR', 'ON')) %>%
  select('fuel_type' = fuel_type_code,
         street_address,
         city,
         state,
         zip,
         latitude,
         longitude)

# Some traditional fuel info
gas <- readxl::read_xlsx('10333_gasoline_stations_year.xlsx')
colnames(gas) <- gas[1,]
gas <- gas[2:52,]

```


# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize Density Map}
# Scrape area of states https://statesymbolsusa.org/symbol-official-item/national-us/uncategorized/states-size  
sq_mile <- read_html(url) %>%
  html_element('table') %>%
  html_table() %>%
  select('state' = X2, 'state_area' = X3) %>%
  slice(-1) %>%
  mutate('state' = 
           countrycode::countrycode(state,
                                    origin = 'state', 
                                    destination = 'abbreviation',
                                    custom_dict = state_dict)) %>%
  add_row(state = 'DC', state_area = '68') %>%
  mutate(state_area = as.numeric(gsub(",", "", state_area)))

# Join to alt_fuel by state
alt_fuel <- left_join(alt_fuel, sq_mile, by = 'state')

# Group and calculate station density
alt_group <- alt_fuel %>%
  group_by(state, state_area) %>%
  summarize(count = n()) %>%
  mutate('station_per_area' = count / state_area) %>%
  arrange(desc(station_per_area))

# Get US state map data
state_map <- map_data("state") 
dc_map <- state_map %>%
  filter(region == 'district of columbia') %>%
  mutate('state' = stringr::str_replace_all(region, 'district of columbia', 'DC')) %>%
  select(!(region | subregion))

# Wrangle to be easy to join to alt_group
state_map <- state_map %>%
  filter(region %notin% 'district of columbia') %>%
  mutate('state' = stringr::str_to_title(region)) %>%
  mutate('state' = countrycode::countrycode(state,
                                    origin = 'state', 
                                    destination = 'abbreviation',
                                    custom_dict = state_dict)) %>%
  select(!(region | subregion)) #%>%
  rbind(dc_map)

# Join to make alt group with shape data  
shape_alt_group <- inner_join(alt_group, state_map, by = 'state')

# Create density plot
light_density <- ggplot() +
  geom_polygon(data = shape_alt_group, 
               aes(x = long, y = lat, group = group,
               fill = station_per_area),
               color = 'black', size = 0.1) +
  scale_fill_continuous(name = ' Stations \n(Per Mile Squared)',
                        low = 'white',
                        high = 'forestgreen',
                        breaks = scales::pretty_breaks()) +
  ggthemes::theme_map() +
  theme(legend.position = "right") +
  labs(title = 'Density of Public Alternative Fuel Stations in the US',
       caption = 'TidyTuedsay | Data: https://afdc.energy.gov/data_download/alt_fuel_stations_format')

# Same graph but for gas stations
gas <- gas %>%
  filter(State %notin% 'Dist. of Col.') %>%
  mutate('state' = stringr::str_replace_all(State, 'New Carolina', 'North Carolina')) %>%
  select(state, stations = '2012') %>%
  mutate(state = countrycode::countrycode(state,
                                  origin = 'state', 
                                  destination = 'abbreviation',
                                  custom_dict = state_dict)) %>%
  left_join(sq_mile, by = 'state') %>%
   mutate(station_area = stations / state_area)


shape_gas <- inner_join(gas, state_map, by = 'state')

gas_map <- shape_gas %>% 
  ggplot() +
  geom_polygon(data = shape_gas, 
               aes(x = long, y = lat, group = group,
               fill = station_area),
               color = 'black', size = 0.1) +
  scale_fill_continuous(name = 'Gas Stations \n(Per Mile Squared)',
                        low = 'white',
                        high = 'red',
                        breaks = scales::pretty_breaks()) +
  ggthemes::theme_map() +
  theme(legend.position = "right") +
  labs(title = 'Density of Public Gas Stations in the US',
       subtitle = 'in the year 2012',
       caption = 'TidyTuedsay | Data: https://afdc.energy.gov')


pdf("compareFuel.pdf")
gridExtra::grid.arrange(light_density, gas_map, nrow=2)
dev.off()
```


```{r round 2}
fuelIcons <- awesomeIconList(
  BD = makeAwesomeIcon(icon = "fa-solid fa-gas-pump",
                       library = "fa",
                       iconColor = "green",
                       markerColor = "green"),
  CNG = makeAwesomeIcon(icon = "fa-solid fa-minimize",
                        library = "fa",
                        iconColor = "green",
                        markerColor = "blue")
)

getColor <- function(alt_fuel) {
  sapply(alt_fuel$fuel_type, function(fuel_type) {
    if(fuel_type == 'CNG') {
      "orange"
    }
    else if(fuel_type == 'BD') {
      "blue"
    }
    else if(fuel_type == 'ELEC'){
      "green"
    }
    else if(fuel_type == 'E85'){
      "yellow"
    }
    else if(fuel_type == 'HY'){
      "purple"
    }
    else if(fuel_type == 'LNG'){
      "cyan"
    }
    else {
      "darkred"
    }
  })
}


alt_fuel %>% 
  filter(fuel_type != 'CNG' & state == 'UT') %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(~longitude, ~latitude, 
                    #icons = ~fuel_type,
                    #color = getColor(),
                    popup = ~paste(fuel_type, street_address, sep = '<br>'))
```
# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(
  filename = "3-1DensityMap.png",
  device = "png")

```
